### 2019-02-05 figures/neighbours.tex

\documentclass[tikz]{standalone}
\usetikzlibrary{arrows.meta,math}
\usepackage{xcolor}
\begin{document}
\begin{tikzpicture}[>=Stealth]
	\node[circle,fill=black!50,draw,label=above right:$b$]{};
	\draw (0,0)--(up:1.5cm);
	\foreach \n/\d/\l/\a/\r in { 
			0/1.4cm/b/10/3cm, 
			1/1.3cm/b'/60/4cm, 
			2/1.2cm/b''/120/3cm ,
			3/1.1cm/b'''/170/3cm ,
			4/1.0cm/b''''/220/3cm 
			} {
		\draw[->] (0,0)--node[above,sloped, very near end]{$r_\n$} (90-\a:\r) 
					node[circle,fill=black!50,draw,label=above right:$\l$]{};
		\draw[->,dashed] (up:\d) arc[start angle=90, delta angle=-\a,
		radius=\d] node[right]{$\beta_\n$};

	};
	\foreach \n/\b/\t in { 0/10/50 , 1/60/60 , 2/120/50 , 3/170/50 , 4/220/150 } {
		\draw[dotted,->] (90-\b:2cm) arc[start angle=90-\b, delta
		angle=-\t,radius=2cm];
		\draw (90-\b-\t/2:2.2cm) node {$\theta_\n$};
	};
\end{tikzpicture}
\end{document}